---
title: "ARIMAX"
author: "tomchaffey"
date: "`r Sys.Date()`"
output: html_document
---
# Utilising ARIMAX modelling

### Loading packages
```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(forecast)
library(xts)
library(dplyr)
library(ggplot2)
library(dygraphs)
library(data.table)
library(fpp2)
library(urca)
```



### Arimax start
We're going to pilot test ARIMAX here, or multifactorial ARIMA. We've already put together some time series for both phytoplankton abundance and river murray discharge. And, attempted ARIMA on phyto abundance to some success. Will have to combine both sets into one df, then go from there. 

We're going to pilot test on Lock 9 with Anabaena. Has a long series of data for both discharge and phyto abundance. 

```{r}
head(anabaena_lock9)
head(LOCK_9)
```

```{r}
anabaena_arimax <- anabaena_lock9 %>%
  select(CELLCOUNT, Date_Sampled)

lock9_arimax <- LOCK_9 %>%
  select(`#Timestamp`, Value)


```

Lock 9 discharge dates all the way back to 1967, we will have to cut down to 1997 (start of Anabaena at lock 9). Importantly, discharge data is taken daily, we should be easily able to cut down to one reading per week, and have two data frames with the same length. 

```{r}
lock9_arimax$`#Timestamp` <-  as.Date(lock9_arimax$`#Timestamp`, tz = "")

str(lock9_arimax)
```

```{r}
lock9_arimax1 <- lock9_arimax %>%
  filter(`#Timestamp` >="1997-11-25") %>%
  filter(`#Timestamp` <="2018-2-20")


lock9_arimax1$`#Timestamp` %>%
  max()

lock9_arimax1$`#Timestamp` %>%
  min()
```

We now have a df for discharge which extend from both the start of our anabaena set to the end of our anabaena set (1997 - 2018). However, our anabaena set includes cellcounts for all reported taxa within the classificiation Anabaena, which range from genus to species, so we often see multiple cell counts on the same day, not sure if this is a problem yet, if necessary, I can only take unique dates from Anabaena to have to df of the same length. 


### Comparative ts plots for both
Given we have ts of equal time length for both, lets take a look and see if theres any correlations. 
```{r}
l9_xts <- xts(lock9_arimax1$Value, order.by = lock9_arimax1$`#Timestamp`)

ab.xts <- xts(anabaena_arimax$CELLCOUNT, order.by = anabaena_arimax$Date_Sampled)

dygraph(l9_xts, main = "Lock 9 Discharge", ylab = "Discharge [cumec]")

dygraph(ab.xts, main = "Anabaena Abundance at Lock 9", ylab = "Cell Count [per mL]")
```
No expert, but if i had to guess, looks like there needs to be a long period of low discharge for there to be a cyaono-bacteria bloom. But thats what modelling is for. 

Lets now subset the Lock 9 discharge to get one reading per week, which we can then combine with our Anabaena df to run ARIMAX. 


```{r}
t1 <- lock9_arimax1 %>%
  mutate(week = floor_date(`#Timestamp`, "week"))

t3 <- t2 %>% 
  group_by(week) %>% 
  summarise_if(is.numeric, mean)
```

We now have week average discharge data for Lock 9, to compare against our weekly data for phytoplankton abundance. Perfect. 

```{r}
lock9.ts <- ts(t2$Value, frequency = 52)
anabaena.ts <- ts(anabaena_arimax$CELLCOUNT)

plot(lock9.ts)
plot(anabaena.ts)
```
Looks alright, continuing. Lets combine. 

Initial problem, while both sets are in Weeks, our anabaena data set extends over multiple taxa, meaning we have a range of Cellcounts of the same day, sometimes 1, sometimes 3. 

Run distinct on anabaena to match data sets. N.B. dont know if this is cool to do or not... 

```{r}
anabaena_arimax1 <- anabaena_arimax %>%
  group_by(Date_Sampled) %>%
  summarise_if(is.numeric, sum)
```
This sums data over the week if there are multiple reported taxa in the same period. 

Problem x: anabaena data is not weekly, didn't realise until.. now. 

Try indexing the sets to match?
```{r}
t3 <- rename(t3, "Date_Sampled" = "week")



```

```{r}
t3 %>% add_column(CELLCOUNT = NA, .after = 2)

anabaena_arimax1 %>% add_column(Value = NA, .after = 1)
```


```{r}
aa <- full_join(t3, anabaena_arimax1)
```




```{r}
discharge.l9 <- lock9_arimax1 %>%
  filter(`#Timestamp` >="1997-11-25") %>%
  filter(`#Timestamp` <="2018-2-20")
```


```{r}
dcl9 <- rename(discharge.l9, "Date_Sampled" = "#Timestamp")
```


```{r}
aa <- full_join(dcl9, anabaena_arimax1)
head(aa)
```


```{r}
aa1 <- aa %>%
  drop_na(CELLCOUNT)

str(aa1)
```

So, here was a roundabout way to solving the problem, rename the dates to match, full_join the data sets, and filter out cellcount NA's. Took a while ,idiot. 
We've now got matching vector for both cellcount and discharge value and can being ARIMAX

### Arimax for real this time

Forecasting with ARIMAX is a bit more involved. We are now using Discharge value to predict or forecast our phytoplankton abundance. We can't just plug it in. We also have to calculate how our xreg (discharge) will change in the future first, and use that within our ARIMAX forecast. In this instance, we will try forecast through 1 year (52 time periods given data is 'mostly' weekly). 

Changing out data into ts objects. 

```{r}
discharge.ts <- aa1 %>%
  select(Value) %>%
  as.ts(start = 1997, frequency = 52)

ab.ts <- aa1 %>%
  select(CELLCOUNT) %>%
  as.ts(start = 1997, frequency = 52)
```


```{r}
fit.aa <- auto.arima(ab.ts,
                     xreg = discharge.ts)

checkresiduals(fit.aa)

```
Calculating x regressors
```{r}
discharge.forecast <- discharge.ts %>%
  mean() %>%
  rep(52)

discharge.forecast # this is the idea, meant to work 
```

```{r}
fcast <- fit.aa %>%
  forecast(xreg = discharge.forecast)

fcast # this is the idea, yet to work 
```

Heres what Rob Hynderman says about forecast with x regressors. You have to create a training model with your xreg, and use it to predict. Will make heads and tails of it soon. 

```{r}
library(forecast)
set.seed(1)
stack_data <- ts(rnorm(100), start = 2000, frequency = 12)
stack_data_xreg <- ts(rnorm(100), start = 2000, frequency = 12)

farma <- function(y, h, xreg) {
  ncol <- NCOL(xreg)
  X <- matrix(xreg[seq_along(y), ], ncol = ncol)
  if (NROW(xreg) < length(y) + h) {
    stop("Not enough xreg data for forecasting")
  }
  newX <- matrix(xreg[length(y) + seq(h), ], ncol = ncol)
  fit <- auto.arima(y, xreg = X)
  forecast(fit, xreg = newX, h = h)
```

