---
title: "Phytoplankton Trend Analysis"
author: "Tom Chaffey"
date: "`r Sys.Date()`"
output: html_document
---

# Phytoplankton trend analysis:

Welcome, have fun watching me do this.

### Relevant packages

Range of packages are needed for this analysis. Both in data tidy-up (see tidyverse/dplyr), time-series forecasting (forecast, tseries, DMwR2), date management (lubridate/linelist) and plotting (ggplot2).

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(DMwR2)
library(forecast)
library(tseries)
library(lubridate)
library(linelist)
library(fpp2)
library(data.table)
library(fable)
```

### Load data

```{r}
Phyto_AWQC <- read.csv(file = 'PMP Data/"Phyto_AWQC.csv')
```

### Data-fixing

While redundant columns have been removed in excel, there is still some work for us to do to get this into a readable format for r. First here being the data. We will fix this first using the r package "lubridate". Dates are thankfully in the same format for both data set, and within the data sets. For now we'll just use the AWQC lab worked data set. 

```{r}
Phyto_AWQC <- Phyto_AWQC_FIXED %>% 
  mutate(Date_Sampled = as.Date(DATESAMPLED, format= "%d/%m/%y")) %>%
  select(-DATESAMPLED)
```

Now R can read the date column successfully, using %y ensures "98" is read as "1998".

We've got multiple NA's among our observations. Recorded as presence/absence for some phytoplankton observations, where there is presence absence, CELLCOUNT is NA and Comments records presence. We will drop these NA's for the time being until I figure out what to do.

```{r}
Phyto_AWQC1 <- Phyto_AWQC %>%
  drop_na(CELLCOUNT)

Phyto2 <- Phyto_AWQC1 %>%
  select(-COMMENTS)
```

```{r}
head(Phyto2)
```

### Exploratory Analysis

We can run some exploratory statistics on what we have so far, paritcularly in regards to the phytoplankton we have amongst this sample.

```{r}
Phyto2 %>%
  group_by(SITENAME) %>%
  summarise(
            unique_taxa = n_distinct(TAXA),
            no_sampling_dates = n_distinct(Date_Sampled))

Phyto2 %>% 
  summarise(no_samplingdates = n_distinct(Date_Sampled),
            no_sites = n_distinct(SITENAME),
            no_taxa = n_distinct(TAXA)
            )
```

15 sites, as expected per the PMP info sheet. 2816 Sampling dates recorded in total. Though broken down into subsequent Sites, we see the most sampling done by far is for Lock 9 and Morgan. Followed far off by Goolwa B, the rest falling \~150 sampling dates. In terms of taxa, lots here, looks like we've got 222 recorded taxa of phytoplankton, this ranges though through species to families, not an exact science. Broken up by Site reveals a bit more consistancy than sampling dates, a rough mean being \~100, tge most at, again, both Lock 9 and Morgan.

Major taxa reported in MDFRC (2011)
 - Total cyanobacteria
 - Anabaena (all?)
 - Aphanizomenon
 - Other Cyanophyceae
 - Aulacoseira granulata
 - Aulacoseira distans
 - Actinastrum 
 - Ankistrodesmus 
 - Cryptohyceae
 Mix of taxa to species, genus and family (And a analysis for total). 
 
 As reported in other analysis, around 70% of taxa are reported as just 1 unit per sampling period. Typically goes unreported (why?).

```{r}
Phyto2 %>%
  group_by(TAXA) %>%
  summarise(count = n())
```

No we can briefly look at some of the taxa we have in this set, there's a lot (222 unique taxa recall), numbers seem to range wildly, as some taxa are recorded to species, others to genus and family. N.B. this doesn't include taxa recorded just as presence/absence, as we removed that, ha.

We can do some exploratory graphing. 

```{r}
Phyto2 %>%
  group_by(SITENAME) %>% 
  filter(TAXA == "Algae - Total") %>%
  sample_n(1000, replace = TRUE) %>%
ggplot(aes(Date_Sampled, CELLCOUNT)) +
  geom_point() +
  scale_y_log10()
```

Here, just grouped by site, and picked a relatively numerous observation (Algae - Total), and added a y log10 scale, since we are dealing with large numbers here, as per the advice of MDFRC (2011). Also took a sample of just 1000 observations for a cleaner view. Definitely some seasonal variation involved here, as to be expected and aligns with MDFRC (2011) findings.

For future, we're going to add a vector for log10 of Cell count.

```{r}
Phyto2$log_cellcount <- log10(Phyto2$CELLCOUNT)
```

### Time series starts

Lets see if we can do some time series forecasting. For this, we're going to be using Total - Algae and Lock 9, taxa and site with lots of data for us to mess around with.

```{r}
algae <- Phyto2 %>%
  filter(SITENAME == "LOCK 9") %>% 
  filter(TAXA == "Algae - Total")

algae$Date_Sampled %>%
  min()

algae$Date_Sampled %>%
  max()

ggplot(algae, aes(Date_Sampled, CELLCOUNT)) +
  geom_point() +
  scale_y_log10() +
  labs( x = "Date", 
        y = "Cell Count (cells/ml)",
        title = "Total Algae at Lock 9")
```

We start at 2006, and end at 2019, nice.

We're now going to just look at Anabaena, one of the most numerous taxa reported in the set.
```{r}
anabaena <- Phyto2 %>%
  filter(grepl("Anabaena", TAXA))

head(anabaena)

anabaena %>%
  group_by(TAXA) %>%
  summarise(count = n())
  
anabaena %>% 
  group_by(UNITS) %>%
  summarise(count = n())

```
So we've got 9 reported taxa for Anabaena. 1 Family, the rest are to species, it also includes different reported in both coiled/straight. We also see that there are 3 different units for reporting this, the dominant, is cells/ml, colonies/ml and mm3/l, lets remove that last two. 

```{r}
anabaena %>%
  filter(!(UNITS == 'colonies/mL')) %>%
  filter(!(UNITS == 'mm3/L')) %>%
  group_by(UNITS) %>%
  summarise(count = n())
```

We've got our df, and our start and end dates. Lets make a time series object in r. In order for our object to not be 15mb large, we're also going to cut out the fat.

N.B Anabaena to be used later. 

```{r}
algae1 <- algae %>%
  select(CELLCOUNT, Date_Sampled)

ts_algae <- ts(algae1, start = 2006,
               frequency = 40
               )
head(ts_algae)
```

```{r}
algae2 <- algae1 %>%
  select(CELLCOUNT)

ts_algae1train <- ts(algae2, start = 2006, end = 2016, frequency = 24)
ts_algae1test <- ts(algae2, start = 2017, end = 2019, frequency = 24)
```

Everyone looks about correct, we now have a ts object in r. Lets split it up into training and test data, for our forecast to us.

```{r}
ts_algae_train <- window(ts_algae, start = 13738, end = 17169, frequency = 40)
ts_algae_test<- window(ts_algae, start = 17176, end = 18128, frequency = 40)
```

Bit finicky, one has to convert dates manually and use them as start/ends.

lets fit baby, we're using triple Z here for our parameters, this means r will auto-assign values based on what it thinks is best.

```{r}
fit1 <- ets(ts_algae1train, model = "ZZZ")

fit1forecast <- forecast(fit1, h = 10)
fit1forecast
```

```{r}
plot(fit1forecast)
```

Lets plot our observed results on top, to see how well this forecast works.

```{r}
ts_algae_auto2 <- ets(as.ts(algae2), model = "ZZZ")
autoplot(ts_algae_auto2)

# dont think this works...
```

This is difficult so we're going to try the same thing using library(fpp2) as recommended by from forecast.

```{r}
ts_algae_auto <- ts(algae2, start = 2006, frequency = 52)

autoplot( ts_algae_auto) +
  labs( x = ("Sampling Date"), y = ("Cell Count (m/l)"), title = ("Lock 9 Total Algae"))

# this looks significantly better i think.
```

Now its fixed, awesome we can get to work. 

```{r}
fit <- Arima(as.ts(algae2), order = c(0,1,5))

checkresiduals(fit)

autoplot(forecast(fit))

#this is what im after, however its missing dates and etc. 
```

```{r}
auto.fit <- auto.arima(as.ts(algae2), lambda = 0)

checkresiduals(auto.fit)
plot(forecast(auto.fit))

#same deal, just using the auto.arima function instead of the standard ARIMA function where parameters must be named, this just automatically fits what it thinks works best. N.B this looks different from an early plot for algae cell count at lock 9, which scaled the y axis, while this remains unscaled. Try forecasting using logged cell counts for better results???

# Need to play around with ARIMA parameters to see if theres a better forecasting model, currently the ts extends over 10 years (for algae), maybe cut down data to a smaller time window for better presentation. 
```
Lets see how this works with Anabaena 

```{r}
anabaena %>% 
  group_by(SITENAME) %>%
  summarise(count = n())

anabaena$Date_Sampled %>%
  max()

anabaena$Date_Sampled %>%
  min()

# Again lock 9 is the most numerous, use it again. 

anabaena_lock9 <- anabaena %>%
  filter(SITENAME =="LOCK 9")

head(anabaena_lock9)
```

Now running an arima 
```{r}
anabaena_ts <- anabaena_lock9 %>%
  select(CELLCOUNT)

anabaena_arima <- auto.arima(as.ts(anabaena_ts), lambda = 0)

checkresiduals(anabaena_arima)

plot(forecast(anabaena_arima))
```


```{r}
anabaena.model <- Arima(
  window(as.ts(anabaena_ts), end = 2012), order =c(0,1,1),
  seasonal = list(order=c(0,1,1), period = 12), lambda = 0)

plot(forecast(anabaena.model, h =52))

# trying to add parameters to hand-done arima to inc accuracy of model, difficult not sure how it works quite yet. 
```
 

### Loops
To extract our major taxa for analysis as per MDFRC (2011), including: 
 - Total cyanobacteria
 - Anabaena (all?)
 - Aphanizomenon
 - Other Cyanophyceae
 - Aulacoseira granulata
 - Aulacoseira distans
 - Actinastrum 
 - Ankistrodesmus
 - Cryptohyceae







