---
title: "Discharge"
author: "tomchaffey"
date: "`r Sys.Date()`"
output: html_document
---

# River Murray Discharge Analysis

This is a part 2 to what we've done prior, we know now that phytoplankton abundance can (somewhat successfully) be modelled via ts in ARIMA. So, we can begin looking at our River murray discharge data. This is only available for some sites along the River murray, so our analysis will start to become more and more limited.

### Loading packages

Same as last time. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(DMwR2)
library(forecast)
library(tseries)
library(lubridate)
library(linelist)
library(fpp2)
library(data.table)
library(fable)
library(dygraphs)
library(xts)
```

### Loading our data

```{r, include=FALSE}
discharge.files <- list.files("Murray_Discharge/")

BALRANALD <- read_csv("Murray_Discharge/BALRANALD.csv")
BURTUNDY<- read_csv("Murray_Discharge/BURTUNDY.csv")
HEYWOODS<- read_csv("Murray_Discharge/HEYWOODS.csv")
LOCK_9<- read_csv("Murray_Discharge/LOCK_9.csv")
MERBEIN<- read_csv("Murray_Discharge/MERBEIN.csv")
MORGAN<- read_csv("Murray_Discharge/MORGAN.csv")
SWAN_HILL<- read_csv("Murray_Discharge/SWAN_HILL.csv")
YARRAWONGA<- read_csv("Murray_Discharge/YARRAWONGA.csv")

```

Mutating files to incl. site name. 
```{r, include=FALSE}
BALRANALD <- BALRANALD %>%
  mutate(Site = "BALRANALD")

BURTUNDY <- BURTUNDY %>%
  mutate(Site = "Burtundy")

HEYWOODS <- HEYWOODS %>%
  mutate(Site = "Heywoods")

LOCK_9 <- LOCK_9 %>%
  mutate(Site = "Lock 9")
  
MERBEIN <- MERBEIN %>%
  mutate(Site = "Merbein")

MORGAN <- MORGAN %>%
  mutate(Site = "Morgan")

SWAN_HILL<- SWAN_HILL %>%
  mutate(Site = "Swan_Hill")

YARRAWONGA <- YARRAWONGA %>%
  mutate(Site = "Yarrawonga")
```

Binding them together for ease of analysis 
```{r, include=FALSE}
discharge_all <- bind_rows(
  BALRANALD, BURTUNDY, HEYWOODS, LOCK_9, MERBEIN, MORGAN, SWAN_HILL, YARRAWONGA
)
```

```{r}
str(discharge_all)
```

### Cleaning Data
```{r}
discharge1 <- discharge_all %>%
  select(-`Quality Code`)

discharge2 <- discharge1 %>%
  select(-'Interpolation Type')

discharge.df <- discharge2 %>%
  drop_na(Value)
```


using the xts function, just found out about it, works like a charm for plotting time series data, better than using dumb dumb ARIMA. 
```{r}
discharge_xts <-  xts(discharge.df$Value, order.by = discharge.df$`#Timestamp`)
```



### Basic time series plots
```{r}
str(discharge.df)
```
```{r}
discharge_Balranald <- discharge.df %>%
  filter(Site =="BALRANALD")

balranald.xts <- xts(discharge_Balranald$Value, order.by = discharge_Balranald$`#Timestamp`)

dygraph(balranald.xts, main = "Balranald Downstream Flow", ylab = "Water Discharge [cumec]")
```


### Combining phytoplankton data and flow data 

We've now got data that is workable for both phyt abundance as a ts and river gauge height (discharge) as a time series. We can now try combine the two (only visually so far, modelling comes after). 

```{r}
lock9.xts <- xts(LOCK_9$Value, order.by = LOCK_9$`#Timestamp`)

dygraph(lock9.xts, main = "Lock 9 Discharge", ylab = "Water Discharge [cumec]")
```
Our lock 9 discharge data extends back to the 1970's which is great. 

Lets look at phyto for lock 9 now. We will use the xts function for this as well so we can combine the two visually. 

```{r}
anabaena_lock9.xts <- xts(anabaena_lock9$CELLCOUNT, order.by = anabaena_lock9$Date_Sampled)

dygraph(anabaena_lock9.xts, main = "Anabaena Cell Count at Lock 9", ylab = "Cell Count (u/mL)")
```

```{r}
for (i in length(unique(discharge_all$Site))){
  xts(discharge_all$Value, order.by = discharge_all$`#Timestamp`) %>%
  dygraph()
}
```



```{r}
for (i in length(unique(discharge_all$Site))){
  xts(discharge_all$Value, order.by = discharge_all$`#Timestamp`) <- paste0(i,"xts")
}
```

